Traditional agroforestry systems in Timor-Leste can store large amounts of carbon in both soil and biomass

### 2024-03-20 manuscript.Rmd
consolidate calculations

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
packages_needed = c("ggplot2", "readxl", "data.table", "knitr")

@@

@@

# for now : download manually from temporary link
# utils::unzip("dataverse_files.zip")
```

@@

# merge plot coordinates with general information (village, AFS)
data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF")])

# simplify site names
data_plot[, village := substr(ID_VILLAGE, 1, 1)]
data_plot[village == "S", village := "Osso Luga"]
data_plot[village == "G", village := "Gariuai"]

## village location: create box around plots
data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(village)]
data_box[village == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
data_box[village == "Osso Luga", w := 0.05]
data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]
```

```{r get-env-data}
# extract elevation and rainfall for the 2 villages
village_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
                                           seq(min(lat), max(lat), 0.001))),
                     .(village)]

colnames(village_env) = c("village", "long", "lat")

# TODO add code to download prec and elev
village_env[, prec := terra::extract(
  terra::rast("D:/maps/CHELSA/CHELSA_bio12_1981-2010_V.2.1.tif"), 
  cbind(long, lat)
)]

village_env[, elev := terra::extract(
  terra::rast("data/srtm_62_14.tif"), 
  cbind(long, lat)
)]

village_summary = village_env[, .(
  coord = paste(round(mean(long),2), round(mean(lat),2), sep = ", "), 
  prec = round(mean(prec)), 
  elev = paste(round(quantile(elev, c(0.025, 0.975))), collapse = " - ")),
  .(village)]
```

The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste. Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r village_summary[village=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r village_summary[village=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r village_summary[village=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r village_summary[village=="Osso Luga", prec]` mm and an altitude ranging `r village_summary[village=="Osso Luga", elev]` m (Fig. \@ref(fig:map-sites)). The average soil properties differed between villages (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture. The two villages also have had different histories. Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2000. During the period of abandonment, the various agroforestry systems in Osso Luga lay fallow.

@@

Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured. Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot: within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured. Height measurements were done with a dendrometer.

Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using inaturalist and Pl@ntnet application, Geoffrey Hull’s monograph (Hull, 2006), the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium). The correspondence between tetum names and scientific names is provided in the supplemental material (Table S2).

@@

In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (see figure S1). Soils were sampled with an auger at 0-10 cm and 20-30 cm. Soils samples were then air-dried and sieved at 2 mm. They were the subject of standard physical and chemical analyses at Cirad laboratories in Montpellier, France. Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.

@@

We estimated tree-level AGC stocks from the measured tree CBH and height using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017]. Each tree was assigned a mean and standard deviation of wood density at the species or genus level when at least one value was available from the Global Wood Density Database [@Zanne2009]. Other taxa were assigned a plot-level average wood density. The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density, taken here as the average wood density in the plots (mean: xxx and standard deviation: xxx). The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2023]. The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.

@@

@@

@@

Soil organic carbon stocks (in Mg.ha-1) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:

@@

@@

```{r agb-calculation}
# get agb data
source("script-agb.R")
```

```{r open-soil-data}
# open organic carbon data
# bulk density in Gariuai according to SoilGrids = 125 cg/cm3
data_soil <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")

# add bulk density and coarse fraction
data_soil2 <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")

# bulk density from direct measurements
data_soil2$BD <- data_soil2$Sample_DRY/data_soil2$V_cylinder

# coarse fraction
data_soil2$CF <- data_soil2$V_rocks_tot/data_soil2$V_cylinder

# change format of depth to correspond to other soil table
data_soil2$SAMPLE_DEPTH <- factor(data_soil2$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
levels(data_soil2$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")

data_soil <- merge(data_soil, data_soil2, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
data.table::setDT(data_soil)
```

```{r add-missing-values}
## add intermediate soil carbon in 10-20 cm depth (for now: no value)
data_soil[, Corg := mean(`Corg\n`, na.rm = TRUE), .(N_PLOT)]
data_soil[!is.na(`Corg\n`), Corg := `Corg\n`]

# replace one missing value of coarse fraction with plot level mean
data_soil[, CF_site := mean(CF, na.rm = TRUE), .(N_PLOT)]
data_soil[is.na(CF), CF := CF_site]
```

```{r total-soc}
# calculate total soil carbon stock per depth
data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]
data_soil[, Cstock :=             ## in Mg/ha
            (h1 - h0) * 1e8 *     ## equivalent volume of a 1 ha soil sample, in cm3/ha
            Corg/100 *            ## carbon organic content in fine soil (unitless)
            (Sample_DRY * 1e-6) * ## mass of the fine fraction of the soil (Mg)
            (1/V_cylinder)        ## 1/volume of the cylinder, in cm-3
]

data_soil_C <-
  data_soil[, .(
    meanSOC = sum(Cstock),
    SC10 = Cstock[SAMPLE_DEPTH == "0-10"],
    clay = mean(as.numeric(CLAY), na.rm = TRUE)
  ), .(N_PLOT)]

setkey(data_soil_C, "N_PLOT")
```

```{r soc-error-propagation}
## error propagation using a Monte Carlo method -------------
data_soil_unc = data_soil[, .(mass = rnorm(1000, Sample_DRY, Sample_DRY*0.1), 
                              Corg = rnorm(1000, Corg, Corg*0.12), 
                              iter = 1:1000), 
                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]

data_soil_unc = data_soil_unc[, .(
  Cstock = sum((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6)), 
  Cstock10 = ((h1 - h0) * 1e8 / 250 * Corg/100 * (mass * 1e-6))[SAMPLE_DEPTH == "0-10"]),
  .(N_PLOT, iter)
]

data_soil_quantile = data_soil_unc[, .(
  lwrSOC = quantile(Cstock, 0.025), 
  medSOC = quantile(Cstock10, 0.5),
  uprSOC = quantile(Cstock, 0.975),
  lwrSOC10 = quantile(Cstock10, 0.025), 
  medSOC10 = quantile(Cstock10, 0.5),
  uprSOC10 = quantile(Cstock10, 0.975)), 
  .(N_PLOT)]
```

```{r combine-agb-soc-data}
# combine AGB and soil C
dataC <- merge(data_agb, data_soil_C, by = "N_PLOT", all=TRUE)

# add soil carbon uncertainty
dataC <- merge(dataC, data_soil_quantile, all=TRUE)

dataC[N_PLOT==6, `:=`(ID_VILLAGE = "GD", ID_AF = "CF")]

dataC[, village := substr(ID_VILLAGE, 1, 1)]
dataC$ID_AF <- factor(dataC$ID_AF, levels = c("CF", "SP", "YA", "HG", "FG"))
```

@@

```{r tab-median, results = "show"}

## bootstrapping

#1. prepare data
data_agc_boot = melt(data_agb_unc, 
                     measure.vars = c("AGC", "BGC"), 
                     id.vars = c("ID_VILLAGE", "N_PLOT", "ID_AF", "iter"))

data_soil_boot = data_soil_unc[, c("N_PLOT", "iter", "Cstock")]
colnames(data_soil_boot)[3] = "value"
data_soil_boot$variable = "SOC"
data_soil_boot$iter = paste0("V", data_soil_boot$iter)

data_soil_boot = merge(data_soil_boot, data_plot[, c("N_PLOT","ID_VILLAGE", "ID_AF")])

data_boot = rbind(data_agc_boot, data_soil_boot[, colnames(data_agc_boot), with = FALSE])

# 2. add total stocks
data_boot_tot = data_boot[, .(value = sum(value), variable = "TOT"), .(ID_VILLAGE, N_PLOT, ID_AF, iter)]
data_boot = rbind(data_boot, data_boot_tot[, colnames(data_boot), with = FALSE])

data_boot[, village := substr(ID_VILLAGE, 1, 1)]

# 3. bootstrap
data_boot_saf = data_boot[, .(value = mean(sample(value, length(value), replace = TRUE))),
                          .(ID_AF, variable, iter)]

# calculate median and 95% confidence interval at site level
data_boot_saf  = data_boot_saf[, .(
  value = paste0(
    signif(mean(value), 3), " (",
    signif(quantile(value, 0.025), 3), "-", 
    signif(quantile(value, 0.975), 3), ")"
  )), 
  .(ID_AF, variable)]

data_boot_saf = dcast(data_boot_saf, ID_AF ~ variable, value = "value")

knitr::kable(data_boot_saf, col.names = c("Agroforestry system", "Aboveground carbon (MgC/ha)", "Belowground carbon (MgC/ha)", "Soil organic carbon (MgC/ha)", "Total carbon (MgC/ha)"))
write.csv(data_boot_saf, file = "tables/C_median_SAF.csv", row.names = FALSE)
```

```{r group-comparison}
dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
## test normality of the data

# log-transform data to get normal distributions (because strictly positive data)
hist((dataC$meanAGC), breaks = 20)
hist((dataC$meanSOC))

# test normality
dataC[, .(
  shap.A = shapiro.test(log(medAGC))$p.value, 
  shap.S = shapiro.test(log(medSOC))$p.value
), .(ID_AF, ID_SUCO)]
# levene test of homoscedasticity
library(car)
dataC[, leveneTest(log(medAGC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]
dataC[, leveneTest(log(medSOC) ~ ID_AF*ID_SUCO)]$`Pr(>F)`[1]

# # pairwise group comparison with Tukey HSD method
# # https://arc.lib.montana.edu/book/statistics-with-r-textbook/item/59
# library(multcomp)
# aovAGC = aov(meanAGC ~ ID_AF, data = dataC)
# tuk <- glht(aovAGC, linfct = mcp(ID_AF = "Tukey"))
# summary(tuk)          # standard display
# tuk.cld <- cld(tuk)   # letter-based display
# 
# plot(tuk.cld)

# # soc
# aovSOC = aov(meanSOC ~ ID_AF, data = dataC)
# tuk <- glht(aovSOC, linfct = mcp(ID_AF = "Tukey"))
# summary(tuk)          # standard display
# tuk.cld <- cld(tuk)   # letter-based display
# 
# opar <- par(mai=c(1,1,1.5,1))
# plot(tuk.cld)

## non parametric test
#updated translation of help page implementation of Nemenyi-Damico-Wolfe-Dunn test
# library(coin)
# is.factor(dataC$ID_AF)
# NDWD <-
#   independence_test(meanAGC ~ ID_AF, data = dataC)
# , distribution = approximate(B = 10000),
# ytrafo = function(data) trafo(data, numeric_trafo = rank_trafo),
# xtrafo = mcp_trafo(iv = "Tukey"))

### global p-value
# print(pvalue(NDWD))
```

```{r tab-anova, results = "show"}
# 2-way anova 
dataC[, ID_SUCO := substr(ID_VILLAGE, 1, 1)]
anov1 = anova(lm(log(medAGC)~ID_SUCO+ID_AF, data = dataC))
anov2 = anova(lm(log(medSOC)~ID_SUCO+ID_AF, data = dataC))
tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
                      ind_var = rep(c("SUCO", "AF"), 2), 
                      p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
)

knitr::kable(tab_anov)
write.csv(tab_anov, file = "tables/anov.csv", row.names = FALSE)

## non parametric 2 way anova: anova on ranks
# anov1 = anova(lm(rank(meanAGC)~ID_SUCO+ID_AF, data = dataC))
# anov2 = anova(lm(rank(meanSOC)~ID_SUCO+ID_AF, data = dataC))
# tab_anov = data.frame(response = rep(c("AGC", "SOC"), each = 2), 
#                       ind_var = rep(c("SUCO", "AF"), 2), 
#                       p = round(c(anov1$`Pr(>F)`[1:2], anov2$`Pr(>F)`[1:2]), 3)
# )

```{r map-sites, fig.height = 6, fig.width = 7}
## create Timor map
bbox <- c(124.5, -9.2, 127.5,  -8.3)
timor_map <- map_data("world", region = "Timor-Leste")

g1 = ggplot(timor_map) + 
  annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
           hjust = 0, vjust = 1, fontface = "bold", size = 6) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
  geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = village)) + 
  labs(x = "", y = "") +
  theme(legend.position = "top", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        panel.background = element_blank())+
  coord_sf(crs = 'WGS84')

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning=FALSE, message=FALSE, results = "hide", tinytex.clean = FALSE)
packages_needed = c("ggplot2", "readxl", "data.table", "knitr", "BIOMASS", "terra", "utils", "car")

@@

**Journal**: Agroforestry Systems

**Max words**: The total length of Original Articles, including figures, tables and references, should not exceed **10,000 words**. The number of figures and tables should not exceed 10 in total.

**Word count**: xxx; **Figure and table count**: 5

@@

# for now : download manually from temporary link
if (!dir.exists("data")) create.dir("data")
# utils::unzip("data/dataverse_files.zip")
```

@@

# merge plot coordinates with general information (site, AFS)
data_plot <- merge(data_coord, data_plot[, c("N_PLOT", "ID_VILLAGE", "ID_AF", "AREA_TOTAL_PARCEL")])

# simplify site names
data_plot[, site := substr(ID_VILLAGE, 1, 1)]
data_plot[site == "S", site := "Osso Luga"]
data_plot[site == "G", site := "Gariuai"]
```

```{r get-env-data}
# precipitation from CHELSA
if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
  url <- "https://os.zhdk.cloud.switch.ch/envicloud/chelsa/chelsa_V2/GLOBAL/climatologies/1981-2010/bio/CHELSA_bio12_1981-2010_V.2.1.tif"
  destfile <- "data/CHELSA_bio12_1981-2010_V.2.1.tif"
  utils::download.file(url, destfile, method = "curl")

  ## crop to reduce memory usage
  box = c(124.5, 127.5, -9.2, -8.3) # Timor Leste
  rst <- terra::rast(destfile)
  rst <- terra::crop(rst, terra::ext(box))
  terra::writeRaster(rst, filename = destfile, overwrite = TRUE) 
}

# 30m resolution SRTM
if (!file.exists("data/CHELSA_bio12_1981-2010_V.2.1.tif")) {
  url <- "https://srtm.csi.cgiar.org/wp-content/uploads/files/srtm_5x5/TIFF/srtm_62_14.zip"
  destfile <- "data/srtm_62_14.zip"
  utils::download.file(url, destfile, method = "curl")
  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
}

# download cwd from Chave et al 2014
if (!file.exists("data/CWD.tif")) {
  url <- "http://chave.ups-tlse.fr/pantropical_allometry/CWD.tif.zip"
  destfile <- "data/CWD.tif.zip"
  utils::download.file(url, destfile, method = "curl")
  utils::unzip(destfile, exdir = "data", overwrite = TRUE)
}
```

```{r extract-env-data}
# extract elevation and rainfall for the 2 sites
site_env = data_plot[, as.list(expand.grid(seq(min(long), max(long), 0.001), 
                                              seq(min(lat), max(lat), 0.001))),
                        .(site)]

colnames(site_env) = c("site", "long", "lat")

# TODO add code to download prec and elev
site_env[, prec := terra::extract(
  terra::rast("data/CHELSA_bio12_1981-2010_V.2.1.tif"), 
  cbind(long, lat)
)]

site_env[, elev := terra::extract(
  terra::rast("data/srtm_62_14.tif"), 
  cbind(long, lat)
)]

site_summary = site_env[, .(
  coord = paste(round(mean(long),2), round(mean(lat),2), sep = ", "), 
  prec = round(mean(prec)), 
  elev = paste(round(quantile(elev, c(0.025, 0.975))), collapse = " - ")),
  .(site)]
```

The research was carried out in the center of the Baucau district, on the north-eastern coast of Timor-Leste. Two communities, or suco (group of villages), were selected to represent two contrasting topo-geographical, climatic and historical cases: (i) Gariuai suco (long-lat coordinates: `r site_summary[site=="Gariuai", coord]`) is located on a plateau of Baucau limestone, with an annual rainfall of `r site_summary[site=="Gariuai", prec]` mm and an altitude ranging 300-750 m; and (ii) the village of Osso Luga in Samalari suco (long-lat coordinates: `r site_summary[site=="Osso Luga", coord]`) is located on the Matebian foothill based on Bobonaro Scaly Clay, with an annual rainfall of `r site_summary[site=="Osso Luga", prec]` mm and an altitude ranging `r site_summary[site=="Osso Luga", elev]` m (Fig. \@ref(fig:map-sites)). The average soil properties differed between sites (Table \@ref(tab:stab-site-soil)), with Gariuai having Clay Loam soil texture compared to Osso Luga having clay soil texture. The two sites also have had different histories. Gariuai has been inhabited for at least 50 years, while Osso Luga was abandoned during the Indonesian occupation, and was only repopulated after 2000. During the period of abandonment, the various agroforestry systems in Osso Luga lay fallow.

@@

Within each sampling plot, all trees with a circumference at breast height (CBH) greater than or equal to 100 cm were identified taxonomically and their CBH and height were measured. Five subplots of 10x10 m (0.01 ha) were established at each corner and in the center of the sampling plot (Fig. \@ref(fig:design)): within these subplots, all trees with a CBH between 30 and 100 cm were taxonomically identified and their CBH and height were measured. Height measurements were done with a dendrometer.

Taxonomic identification was carried out using tetum vernacular names, which were then linked to scientific names using iNaturalist and Pl\@ntnet application, Geoffrey Hull’s monograph [@Hull2006], the data bases of “the useful tropical plants” (https://tropical.theferns.info/) as well as the native plants websites of the NT herbarium from the Australian government (https://nt.gov.au/environment/native-plants/native-plants-and-nt-herbarium). The correspondence between tetum names and scientific names is provided in the supplemental material (Table \@ref(tab:tetum-names)).

@@

In each of the 30 sampling plots, a composite of 3 samples were collected following a transect to cover the plot’s inner-variability (Fig. \@ref(fig:design)). Soils were sampled with an auger at 0-10 cm and 20-30 cm. Soils samples were then air-dried and sieved at 2 mm. They were the subject of standard physical and chemical analyses at Cirad laboratories in Montpellier, France. Among these analyses, total organic carbon was determined by dry combustion and measured with a thermal conductive detector.

@@

```{r agc-estimation}
### 1. Open data

# Open large tree data (> 100 cm circumference)
data_large <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV >100CM")

# Open small tree data (30-100 cm circumference)
data_small <- readxl::read_excel("data/DATA_BIOMASS_TL_SHORT.xlsx", "TREE INV 30-100CM")

### 2. Combine tables

# create a "N_Q_ID" column in data_large, with corresponding quadrat number =
# "TOT" (entire plot)
data_large$N_Q_ID <- paste(data_large$N_PLOT, "TOT", sep = "_")

# common columns to be kept
col_keep <- c("N_PLOT", "N_Q_ID", "SCIENTIFIC_NAME", "DIAM_130_CM", "HEIGHT")

# stack tree tables
data_tot <- rbind(data_large[, col_keep], data_small[, col_keep])

# merge with plot information
data_tot <- merge(data_tot, data_plot, by = "N_PLOT")

# convert data_tot into data.table 
data.table::setDT(data_tot)

### 3. Clean tables

# remove NA or <30 cm circumference values or < 100 cm outside subplots
data_tot <- subset(data_tot, 
                   !is.na(DIAM_130_CM) & 
                     DIAM_130_CM >= 30/pi &
                     (DIAM_130_CM >= 100/pi | grepl("Q", N_Q_ID)))

# split scientific names into genus and species
names <- data.table::tstrsplit(data_tot$SCIENTIFIC_NAME, " ")

# get genera
data_tot$GENUS <- names[[1]]

# get species
data_tot$SPECIES <- names[[2]]

# Timoneus should be Timonius
data_tot$GENUS[!is.na(data_tot$GENUS) & data_tot$GENUS == "Timoneus"] <- "Timonius"

# remove one tree with DBH = 0
data_tot <- subset(data_tot, DIAM_130_CM > 0)

### 4. Get values for Chave et al., 2014 equation

# get wood density
data_wd <- BIOMASS::getWoodDensity(genus = data_tot$GENUS, species = data_tot$SPECIES)

# 15% coefficient of variation on height
cvH <- 0.15

### estimate AGC
data_sim_agc <- data.table(
  data_tot,
  BIOMASS::AGBmonteCarlo(
    D = data_tot$DIAM_130_CM, 
    WD = data_wd$meanWD, 
    errWD = data_wd$sdWD, 
    H = data_tot$HEIGHT, 
    errH = data_tot$HEIGHT*cvH, 
    Dpropag = "chave2004", 
    n = 1000,
    Carbon = TRUE
  )$AGC_simu
) |> 
  melt(
    measure.vars = paste0("V", 1:1000),
    variable.name = "iter",
    value.name =  "AGC"
  )

### 5. Apply a different allometry for palms
palm_genera <- c("Cocos", "Areca", "Arenga", "Corypha", "Borassus")

## height in cm, with a conversion factor between 0.8 and 1
data_sim_agc[GENUS %in% palm_genera, 
             HEIGHT := rnorm(1, HEIGHT, HEIGHT*cvH)*
               runif(1, .8, 1)* 100]

## add WD data
data_wd_palms = unique(data_wd[data_wd$genus %in% palm_genera, c("genus", "meanWD", "sdWD")])
data_sim_agc = merge(data_sim_agc, data_wd_palms, all.x = TRUE, by.x = "GENUS", by.y = "genus")
data_sim_agc[GENUS %in% palm_genera, WD := rnorm(1, meanWD,sdWD)]

data_sim_agc[GENUS %in% palm_genera, 
             AGC := WD * HEIGHT * (DIAM_130_CM/2)^2 * pi * 
               0.89 * # volumetric shrinkage
               1e-6  # convert from g to tons
]
```

We estimated tree-level AGC stocks from the measured tree CBH and height using the pantropical allometric equation from @Chave2014 and the R package BIOMASS [@Rejou-Mechain2017]. Each tree was assigned a mean and standard deviation of wood density at the species or genus level when at least one value was available from the Global Wood Density Database [@Zanne2009]. Other taxa were assigned a plot-level average wood density. The biomass of the palms was estimated by multiplying their estimated volume by their estimated wood density. The volume of the palm trunks was estimated as the volume of a cylinder, using the height and diameter measured in situ, with a volumetric shrinkage of 11% [@WD2023]. The height of the palms was measured above the leaves and was therefore multiplied by a conversion factor of 0.9 to obtain the trunk height.

@@

```{r bgc-estimation}
# root-to-shoot ratio allometry from Ledo et al., 2018
# use CWD values from Chave et al., 2014
data_plot[, cwd := terra::extract(terra::rast("data/CWD.tif"), cbind(long, lat))]
data_sim_agc <- merge(data_sim_agc, data_plot[, c("N_PLOT", "cwd")], by = "N_PLOT")

# add uncertainty on R:S ratio? 
data_sim_agc[, RS := exp(- 1.2312 - 0.0215 * DIAM_130_CM + 0.0002 * DIAM_130_CM^2 - 0.0007 * cwd)]
data_sim_agc[, BGC := RS * AGC]
```

@@

```{r aggregate-biomass-data}
# replace with area = 0.05 for quadrats (5 quadrats of 0.01 ha each)
data_sim_agc[, area := ifelse(grepl("Q", N_Q_ID), 0.01, AREA_TOTAL_PARCEL)]

# melt data frame
data_sim <- melt(data_sim_agc, id.vars = c("N_PLOT", "N_Q_ID", "ID_VILLAGE", "ID_AF", "iter", "area"), measure.vars = c("AGC", "BGC"))

# estimate biomass per quadrat and plot
data_sim <- data_sim[, .(value = sum(value/area)), .(N_PLOT, N_Q_ID, ID_VILLAGE, ID_AF, iter, variable)]

## bootstrap AGC and BGC values in small plots
data_sim = 
  merge(
    data_sim[!grepl("Q", N_Q_ID)], 
    data_sim[grepl("Q", N_Q_ID), .(
      valueSmall = ifelse(
        length(value) == 0,   ## some quadrats have zero trees: complete their values
        0, 
        mean(sample(c(value, rep(0, 5 - length(value))), replace = TRUE))
      )), .(N_PLOT, variable, iter)], 
    by = c("N_PLOT", "variable", "iter")
  )

data_sim[, value := value + valueSmall]
```

@@

Soil organic carbon stocks (in Mg.ha$^{-1}$) were then calculated using the method recommended by @Poeplau2017 (M4), which correctly takes into account the coarse fraction of the soil:

@@

```{r open-soil-data}
# open soil lab data
data_soil_lab <- readxl::read_excel("data/RESUME_DATA_SOIL_SHORT_TL.xls", sheet = "RESUME-LAB ")

# open soil field measurements
data_soil_field <- readxl::read_excel("data/RAW_DATA_SOIL_LARGE_TL.xls", sheet = "SAMPLING_MEASURE")

# change format of depth to correspond to other soil table
data_soil_field$SAMPLE_DEPTH <- factor(data_soil_field$DEPTH_SAMPLE_C, levels = c("2_7", "12_17", "22_27"))
levels(data_soil_field$SAMPLE_DEPTH) <- c("0-10", "10-20", "20-30")

data_soil <- merge(data_soil_lab, data_soil_field, by = c("N_PLOT", "SAMPLE_DEPTH"), all=TRUE)
data.table::setDT(data_soil)
```

```{r add-missing-values}
## add intermediate soil carbon in 10-20 cm depth
data_soil[, Corg := `Corg\n`]
data_soil[, muCorg := mean(Corg, na.rm = TRUE), .(N_PLOT)]
data_soil[is.na(Corg), Corg := muCorg]
```

```{r soc-error-propagation}
data_soil[, c("h0", "h1") := lapply(data.table::tstrsplit(SAMPLE_DEPTH, "-"), as.numeric)]

## error propagation using a Monte Carlo method -------------
data_sim_soil = data_soil[, .(mfine = rnorm(1000, Sample_DRY, Sample_DRY*0.1), # 10% error on fine soil mass
                              Corg = rnorm(1000, Corg, Corg*0.12), # 12% error on C content measurement
                              iter = paste0("V", 1:1000)), 
                          .(N_PLOT, SAMPLE_DEPTH, h1, h0)]

data_sim_soil[, Cstock :=             ## in Mg/ha
                (h1 - h0) * 1e8 *     ## equivalent volume of a 1 ha soil sample, in cm3/ha
                Corg/100 *            ## carbon organic content in fine soil (unitless)
                (mfine * 1e-6) * ## mass of the fine fraction of the soil (Mg)
                (1/250)        ## 1/volume of the cylinder, in cm-3
]

# sum over depths + at 0-10 cm depth
data_sim_soil = data_sim_soil[, .(
  SOC = sum(Cstock), 
  SOC10 = Cstock[SAMPLE_DEPTH == "0-10"]),
  .(N_PLOT, iter)
]

## melt data 
data_sim_soil = melt(data_sim_soil, id.vars = c("N_PLOT", "iter"))
```

```{r merge-all}
## merge with biomass data
data_sim = rbind(data_sim[, colnames(data_sim_soil), with = FALSE], data_sim_soil) |> 
  merge(data_plot[, c("N_PLOT", "site", "ID_AF")], by = "N_PLOT")

# add total carbon stocks
data_sim = rbind(data_sim, data_sim[, .(value = sum(value[variable != "SOC10"]), variable = "TOT"), .(N_PLOT, site, ID_AF, iter)])
```

@@

```{r boot-results}
# plot level statistics
dataC_plot = data_sim[, .(
  mean = mean(value),
  lwr = quantile(value, 0.025),
  med = quantile(value, 0.5),
  upr= quantile(value, 0.975)
), .(plot = N_PLOT, site, afs = ID_AF, variable)]

# AFS level statistics (with bootstrapping)
dataC_afs = data_sim[, .(
  mean = mean(value),
  meanBs = mean(sample(value, replace = TRUE))
), .(afs = ID_AF, variable, iter)][, .(
  mean = mean(mean),        
  lwr_mean = quantile(meanBs, 0.025),   
  upr_mean = quantile(meanBs, 0.975)
), .(afs, variable)]

# site level statistics (with bootstrapping)
dataC_site = data_sim[, .(
  mean = mean(value),
  meanBs = mean(sample(value, replace = TRUE))
), .(site, variable, iter)][, .(
  mean = mean(mean),        
  lwr_mean = quantile(meanBs, 0.025),   
  upr_mean = quantile(meanBs, 0.975)
), .(site, variable)]
```

```{r group-comparison}
# shapiro wilk test - normality
dataC_plot[variable %in% c("AGC", "SOC"), .(
  shap = shapiro.test(log(mean))$p.value
), .(afs, variable)]
# levene test of homoscedasticity
dataC_plot[variable %in% c("AGC", "SOC"), 
           leveneTest(log(mean) ~ afs*site)$`Pr(>F)`[1], .(variable)]

# 2-way anova 
anova_table = 
  dataC_plot[variable %in% c("AGC", "SOC"), 
           .(pvalue = round(anova(lm(log(mean) ~ site + afs))$`Pr(>F)`[1:2], 3)), 
                   .(variable)]
anova_table$factor = rep(c("site", "afs"), 2)
```

@@

```{r summary, results = "show"}
# calculate median and 95% confidence interval at site level
dataC_afs[, value := paste0(
  signif(mean, 3), " (",
  signif(lwr_mean, 3), "-", 
  signif(upr_mean, 3), ")")] |> 
  subset(variable != "SOC10") |> 
  dcast(afs ~ variable, value.var = "value") |> 
  knitr::kable(col.names = c("Agroforestry system", 
                             "Aboveground carbon (MgC/ha)", 
                             "Belowground carbon (MgC/ha)", 
                             "Soil organic carbon (MgC/ha)", 
                             "Total carbon (MgC/ha)")) 
```

```{r anova, results = "show"}
knitr::kable(anova_table[, c("variable", "factor", "pvalue")])
```

```{r map-sites, fig.height = 6, fig.width = 7}

## site location: create box around plots
data_box = data_plot[, .(xc = mean(long), yc = mean(lat)), .(site)]
data_box[site == "Gariuai", `:=`(w = 0.1, xc = xc + 0.001)]
data_box[site == "Osso Luga", w := 0.05]
data_box[, `:=`(x1 = xc - w/2, x2 = xc + w/2, y1 = yc - w/2, y2 = yc + w/2)]

## create Timor map
bbox <- c(124.5, -9.2, 127.5,  -8.3)
timor_map <- map_data("world", region = "Timor-Leste")

g1 = ggplot(timor_map) + 
  annotate(geom = "text", label = "a", x = -Inf, y = Inf, 
           hjust = 0, vjust = 1, fontface = "bold", size = 6) + 
  geom_polygon(aes(x = long, y = lat, group = group), fill="lightgray", colour = "grey") +
  geom_tile(data = data_box, aes(x = xc, y = yc, width = w, height = w), fill = NA, color = 1, linewidth = 0.5) +
  ggrepel::geom_text_repel(data = data_box, aes(x = xc, y = yc, label = site)) + 
  labs(x = "", y = "") +
  theme(legend.position = "top", axis.title = element_blank(), 
        axis.ticks = element_blank(), axis.text = element_blank(), 
        panel.background = element_blank())+
  coord_sf(crs = 'WGS84')

# get elevation information
elev = as.data.frame(terra::rast("data/srtm_62_14.tif"), xy = TRUE)
elevS <- subset(elev, x > data_box[site == "Osso Luga"]$x1 & 
                  x < data_box[site == "Osso Luga"]$x2 & 
                  y > data_box[site == "Osso Luga"]$y1 & 
                  y < data_box[site == "Osso Luga"]$y2)
elevS$site = "Osso Luga"
elevG <- subset(elev, x > data_box[site == "Gariuai"]$x1 & 
                  x < data_box[site == "Gariuai"]$x2 & 
                  y > data_box[site == "Gariuai"]$y1 & 
                  y < data_box[site == "Gariuai"]$y2)
elevG$site = "Gariuai"
